raptr: Representative and Adequate Prioritization Toolkit in R

### 2017-06-12 code/rmarkdown/article.Rmd
assorted improvements based on reviewer's suggestions

```{r, include = FALSE}
## initialization
# load .rda file
try(session::restore.session("data/final/results.rda"))
```

Abstract word count: 341 / 350 \newline
Total word count: 6931 / 7000 \newline
Number of references: 49 \newline
Number of figures: 8 \newline
Number of tables: 0 \newline
Number of textboxes: 0 \newline
Number of equations: 4 \newline
Short running title: \texttt{raptr} R package \newline
Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr  \newline
Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas  \newline

@@

1. An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity. To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved. One way to conserve such processes at the feature level (eg. species, ecosystem) is to preserve a sample of the feature (eg. individuals, areas) that is representative of the intrinsic or extrinsic physical attributes that underpin the process of interest. For example, by conserving a sample of populations with local adaptations--physical attributes associated with adaptation--that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost. Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature. Little attention has been directed towards capturing a representative sample of the variation within each feature.

2. To address this issue, we developed the \texttt{raptr R} package to help guide reserve selection. Users set "amount targets"--similar to conventional methods--to ensure that solutions secure a sufficient proportion of area or habitat for each feature. Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg. environmental or genetic variation). We demonstrate the functionality of this package using simulations and two case studies. In these studies, we generated solutions using amount targets--representing conventional methods--and compared them with solutions generated using amount and space targets.

@@

4. The \texttt{raptr R} package provides a toolkit for making prioritizations that secure an adequate and representative sample of each feature. By using solutions that secure a representative sample of each feature, prioritizations may have a greater chance of achieving long-term biodiversity persistence.

@@

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities, and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using this method is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490]. This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg. @r295].

To overcome this limitation, Faith and Walker [-@r480] developed a decision support tool (\texttt{DIVERSITY}) to identify prioritizations that secure a representative sample of the environmental variation found across a study area. Although this method was originally pitched as an alternative to species-based conservation planning [@r481; @r218; @r203], it can also identify solutions that conserve a representative sample of variation across a single species' range. Recent work has built on this reserve selection method by solving problems with more advanced optimization algorithms [@r482]. However, existing \texttt{DIVERSITY}-based methods have limited utility for species-based conservation planning, in part, because they cannot accommodate multiple biological features and do not provide options to specify specific targets pertaining to the amount of habitat required to conserve species (unlike \texttt{Marxan} and \texttt{Zonation}). As a consequence, such methods are not useful for most conservation planning exercises [@r255].

Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of the variation in each feature. To begin to fill this gap, we unite the ideas underpinning \texttt{DIVERSITY} and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes. Here, we aim to provide an in-depth understanding of this \texttt{R} package and explore its functionality.

# Methods
## PROBLEM FORMULATION
Biodiversity features are defined as the entities that the prioritization is required to preserve (eg. species, ecosystems). Here, attributes are defined as the variation across the features' ranges that the prioritization is required to sample. These attributes can be intrinsic (eg. genetic or phenotypic) or extrinsic (eg. environmental conditions) to the feature. There should, however, be a reasonable underlying hypothesis that relates the attribute to the biodiversity processes that needs to be conserved.

A set of attributes is conceptualized as a space (termed an "attribute space"). For example, a decision maker may require a prioritization that represents populations along climatic gradients. To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this climatic space. Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space. By associating the planning units with climatic data, and calculating a descriptive statistic for each planning unit (eg. mean), they can be mapped from geographic space to this climatic attribute space. In addition to points representing planning units, attribute spaces also contain demand points.

Demand points [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets). For instance, by siting demand points throughout an attribute space, planners can obtain solutions that conserve a representative sample of an attribute space. Alternatively, by siting demand points in specific regions of an attribute space, planners can obtain solutions that target specific samples of an attribute space. For a given set of demand points, the shorter the distances between the demand points and the planning units selected for prioritization; the better a solution is at securing the desired variation in the attribute space. In any attribute space there may exist points that are impossible (eg. mean annual rainfall -5 mm), or do not occur in the study area (eg. mean annual temperature 30$^{\circ}$C in Antarctica). Additionally, there may be some regions that are desirable for some features and undesirable for others (eg. conditions known to be outside the physiological tolerance of certain species). Thus a different set of demand points may be appropriate for different attribute spaces and features.

To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations. Since we can only afford to preserve three of the four populations, our objective is to conserve the most representative sample possible. To achieve this goal, we obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional climatic attribute space. Next, we generated demand points as equidistant points inside this space. By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify the optimal solution (Fig. 1). We can see that a solution that prioritizes both populations _A_ and _B_, effectively "doubles-up" on the same climatic characteristics and constitutes considerable redundancy if both selected. Instead, a more representative sample of the intra-specific variation could be captured by securing populations _A_ (or _B_), _C_, and _D_. However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, we could site demand points in environmental conditions with temperatures over 30 $^{\circ}$C (ie. the top two rows of demand points in Fig. 1). Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized. Since demand points can be sited and weighted in any configuration, they provide a flexible means to guide the reserve selection process.

The \texttt{raptr R} package utilizes two novel formulations of the reserve selection problem to generate prioritizations. These formulations are based on elements of the \texttt{Marxan}, \texttt{DIVERSITY}, and uncapacitated facility location problems [@r451]. Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations. The difference between the two formulations is that the reliable formulation explicitly accommodates uncertainty in the probability of features occupying the planning units when calculating how well a given solution samples a feature's attribute space. For brevity, we will define the simpler formulation--the unreliable formulation--below and define the more complex version--the reliable formulation--in Appendix S1. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of sets will be denoted using the same symbol used to denote the set.

@@

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points. Demand points with greater weight $\lambda_{fsi}$ are more important, so solutions that select planning units close to highly weighted demand points will be more favorable. As a consequence, the decision maker will need to choose an appropriate weighting for each demand point.

Targets are used to ensure that prioritizations sufficiently conserve each feature. Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}). Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$. Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved. Space-based targets are expressed as proportions--instead of a sum of weighted distances--by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid. This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means clustering analyses from the within sums of squares and total sums of squares [page 106; @r491]. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie. a $j \in J$ where $X_j = 1$). The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$). Generally--unless near zero space targets are used so that the problem is effectively unconstrained by the target--demand points are represented by their closest selected planning units. In poorer quality solutions, demand points will be represented by planning units that have larger distances between them. As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.

@@

The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is. Constraints (eqn 3b) ensure that all the amount-based targets are met. Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space. For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}$). This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$). The resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis. The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target. Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point. Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units. Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.

## Optimization
Although the reserve selection problems presented here are non-linear (see Appendix S1 for the reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16]. The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}). Presently, academics can obtain a license at no cost.

@@

We generated four solutions for each species. First, to represent solutions generated using conventional methods (eg. \texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets. Second, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species (hereafter referred to as geographic spread targets). Third, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`. Finally, to illustrate the combined effects of using amount and space targets as well as penalties for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % geographic spread targets, and boundary length modifiers of `r sim_parameters$blm`.

@@

The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units near the southern end of study area (Fig. 2a). For this species, all solutions containing this number of planning units are optimal when using a `r sim_parameters$amount_target*100` % amount target. We obtained this particular solution due to artifacts in the solver (eg. seed for the random number generator, order of variables in the problem matrix). This solution captured a small amount of the geographic spread of the species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled). In fact, the coverage was so poor that the proportion was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution.

By explicitly targeting representativeness, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3b). Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented. However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (Fig. 3d). This solution prioritized the same number of planning units as the previous solutions even though it is far superior. As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.

@@

The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core distribution (Fig. 3e). While this strategy is cost-effective for protecting habitat, it was ineffective for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any planning units along the periphery of the species' distribution. By using amount and space targets, we were able to conserve an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3f). Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (Fig. 3h). However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions. These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.

@@

The solution generated for the bimodally distributed species using just an amount target conserved individuals belonging to one of the two populations (Fig. 3i). As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled). The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount target\nand BLM")$space_held, 2)` % sampled; Fig. 3k). However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % geographic spread secured; Fig. 3l). This finding suggests that species with large intra-specific variation could benefit the most from prioritizations generated using space-based targets.

>>>>

```{r, include = FALSE}
## initialization
# load .rda file
try(session::restore.session("data/final/results.rda"), silent = TRUE)
```

Abstract word count: 346 / 350 \newline
Total word count: ??? / 7000 \newline
Number of references: 49 \newline
Number of figures: 8 \newline
Number of tables: 0 \newline
Number of textboxes: 0 \newline
Number of equations: 4 \newline
Short running title: \texttt{raptr} R package \newline
Short abstract: Prioritize biodiversity processes and patterns using raptr https://github.com/jeffreyhanson/raptr  \newline
Keywords: conservation, biodiversity, mixed integer linear programming, optimization, protected areas  \newline

@@

1. An underlying aim in conservation planning is to maximize the long-term persistence of biodiversity. To fulfill this aim, the ecological and evolutionary processes that sustain biodiversity must be preserved. One way to conserve such processes at the feature level (eg. species, ecosystem) is to preserve a sample of the feature (eg. individuals, areas) that is representative of the intrinsic or extrinsic physical attributes that underpin the process of interest. For example, by conserving a sample of populations with local adaptations---physical attributes associated with adaptation---that is representative of the range of adaptations found in the species, protected areas can maintain adaptive processes by ensuring these adaptations are not lost. Despite this, current reserve selection methods overwhelmingly focus on securing an adequate amount of area or habitat for each feature. Little attention has been directed towards capturing a representative sample of the variation within each feature.

2. To address this issue, we developed the \texttt{raptr R} package to help guide reserve selection. Users set "amount targets"---similar to conventional methods---to ensure that solutions secure a sufficient proportion of area or habitat for each feature. Additionally, users set "space targets" to secure a representative sample of variation in ecologically or evolutionary relevant attributes (eg. environmental or genetic variation). We demonstrate the functionality of this package using simulations and two case studies. In these studies, we generated solutions using amount targets---representing conventional methods---and compared them with solutions generated using amount and space targets.

@@

4. The \texttt{raptr R} package provides a toolkit for making prioritizations that secure an adequate and representative sample of variation within each feature. By using solutions that secure a representative sample of each feature, prioritizations may have a greater chance of achieving long-term biodiversity persistence.

@@

Today, the most widely used conservation planning tools focus on biodiversity patterns [\texttt{Zonation} and \texttt{Marxan}; @r429; @r22]. Decision makers can use these tools to obtain solutions that secure a proportion of the geographic range of each biological feature (populations, species, or ecosystems) of interest, by setting targets. One method to incorporate data on the ecological and evolutionary processes that sustain such features is to conserve a representative sample of the internal or external physical attributes underpinning these processes. To achieve this, current methods typically involve partitioning each feature into sub-groups based on an attribute variable that relates to a biodiversity process of interest [@r9; @r247; @r448] or phylogenetic trees [@r497]. For instance, by dividing species distributions into sub-groups according to habitat discontinuities and ensuring that each sub-group is represented in the solution, conservation planners can obtain prioritizations that promote adaptive processes [@r2]. However, using these methods is challenging because biodiversity often cannot be divided into operational groups without information loss [@r489; @r218], and because the number of groups will substantially alter solutions [@r490]. This limitation has been known for quite some time, and dates back to some of the earliest reserve selection methods [eg. @r295].

To overcome this limitation, Faith and Walker (1996) developed the "environmental diversity" (ED) reserve selection framework to conserve a representative sample of the environmental variation found across a study area. Drawing inspiration from the _p_-median problem [@r498], this method used Euclidean distances to express differences among candidate sites and maximize the representativeness of the solution [@r218; @r203; [@r204]. Recent work has built on this method using more advanced optimization algorithms [@r482]. However, existing environmental diversity based methods are not often used in conservation planning. This is, in part, because they seek to find the most representative solution that is within a budget, unlike the more commonly used decision support tools (eg. \texttt{Marxan}) that aim to find the cheapest solution subject to a set of targets (or constraints).

Conservation planners lack a decision support tool that lets them set explicit targets to obtain solutions that (1) secure an adequate amount of habitat for each feature and (2) a representative sample of the variation in each feature. To begin to fill this gap, we unite the ideas underpinning environmental diversity and \texttt{Marxan} into new formulations of the reserve selection problem, and implement them in the \texttt{raptr R} package. This \texttt{R} package provides decision makers with the tools to generate prioritizations based on data that relate to biodiversity patterns and processes. Here, we aim to provide an in-depth understanding of this \texttt{R} package and explore its functionality.

# Methods
## PROBLEM FORMULATION
Biodiversity features are defined as the entities that the prioritization is required to preserve (eg. species, ecosystems). Each biodiversity feature may have one or more "attributes" which vary across its geographic range as a result of ecological or evolutionary processes. These attributes can be intrinsic (eg. genetic or phenotypic) or extrinsic (eg. environmental conditions) to the feature. The overarching goal of the prioritization is to conserve these biodiversity processes, and by representing the variation in these species-specific attributes, the prioritization can make some progress towards achieving this. Thus there should be a reasonable underlying hypothesis that relates the attributes used to inform the reserve selection processes to the biodiversity process that needs to be conserved.

The variation in the attributes of a feature can be represented using an _n_-dimensional "attribute space". For example, a decision maker may require a prioritization that represents populations along climatic gradients. To achieve this, the decision maker might use a "climatic" attribute space with dimensions relating to mean annual temperature ($^{\circ}$C) and precipitation (mm). Any given combination of temperature and precipitation may be conceived as a point in this climatic space. Although they exist as polygons, for simplicity, each planning unit may be thought to exist as a single point inside a given attribute space. By associating the planning units with climatic data, and calculating a descriptive statistic for each planning unit (eg. mean), they can be mapped from geographic space to this climatic attribute space. In addition to points representing planning units, attribute spaces also contain points that describe which attributes should be conserved.

"Demand points" [@r480; @r218; @r203] are designated by the decision maker to indicate regions of the attribute space that they wish to represent in the prioritization (see below for discussion on generating demand points for real-world data sets). For instance, by siting demand points throughout an attribute space, planners can obtain solutions that better capture existing variation. Alternatively, by siting demand points in specific regions of an attribute space, planners can obtain solutions that target specific samples of an attribute space. For a given set of demand points, the shorter the sum of the distances between the demand points and the planning units selected for prioritization: the better a solution is at securing the desired variation in the attribute space. Demand points are used heavily in the _p_-median and facility location problems [reviewed in @r498]. Additionally, different features and attribute spaces may require different demand points. For instance, points that may be valid in one attribute space (eg. mean annual temperature -5 ${\circ}$C) may be invalid for another attribute space (eg. mean annual rainfall -5 mm). Additionally, there may be some regions that are desirable for some features and undesirable for others (eg. conditions known to be outside the physiological tolerance of certain species).

To illustrate these concepts, consider the following example: we wish to develop a prioritization for a single species that has four populations. Since we can only afford to preserve three of the four populations, our objective is to conserve the most representative sample possible. To achieve this goal, we obtained annual rainfall (mm) and temperature ($^{\circ}$C) data at the location of each population. We used this data to construct a two-dimensional climatic attribute space. Next, we generated demand points as equidistant points inside this space. By comparing the distribution of the demand points to the distribution of the populations in the attribute space, we can identify the optimal solution (Fig. 1). We can see that a solution that prioritizes both populations _A_ and _B_ constitutes considerable redundancy by selecting populations that inhabit similar conditions. Instead, a more representative sample of the intra-specific variation could be captured by securing populations _A_ (or _B_), _C_, and _D_. However, if the goal of the prioritization was to preserve populations living in warmer temperatures, then instead of siting demand points across the full range of conditions, we could site demand points in environmental conditions with temperatures over 30 $^{\circ}$C (ie. the top two rows of demand points in Fig. 1). Given this new set of demand points, populations _A_, _B_, and _C_ would be prioritized. Since demand points can be sited and weighted in any configuration, they provide a flexible means to guide the reserve selection process.

The \texttt{raptr R} package utilizes two novel formulations of the reserve selection problem to generate prioritizations. These formulations are based on ideas that underpin the \texttt{Marxan}, environmental diversity, and uncapacitated facility location problems [@r498]. Since they are based on the unreliable and reliable facility location problems [@r16], the formulations are hereafter referred to as the "unreliable" and "reliable" formulations. The difference between the two formulations is that the reliable formulation explicitly accommodates uncertainty in the probability of features occupying the planning units when calculating how well a given solution samples a feature's attribute space. For brevity, we will define the simpler formulation---the unreliable formulation---below and define the more complex version---the reliable formulation---in Appendix S1. All mathematical terms defined hereafter are described in Table S1. For convenience, the cardinality of a given set will be denoted using the same symbol used to denote the set.

@@

Let $S$ denote a set of attribute spaces (indexed by $s$). Each $j \in J$ is associated with spatially explicit data that represent coordinates for each attribute space $s \in S$. Let $I_{fsi}$ denote a set of demand points (indexed by $i$) for each feature $f \in F$ and each attribute space $s \in S$. Let $\lambda_{fsi}$ denote the weighting for each demand point $i \in I$, $f \in F$ and $s \in S$. Let $d_{fsij}$ denote the distance between each demand point $i \in I$ and each planning unit $j \in J$ for each feature $f \in F$ and attribute space $s \in S$. To describe the inherent variation in the distribution of demand points for feature $f$ and space $s$, let $\delta_{fsi}$ denote the distance between each demand point $i \in I$ and the centroid of the demand points.

Demand points with greater weight $\lambda_{fsi}$ are more important, and so solutions may need to select planning units closer to highly weighted demand points to fulfill space targets. Since increasing the number of demand points in the problem also increases the time required to solve it, conservation planners may find it more useful to increase of the weighting of demand points in important regions of an attribute space rather than increasing the number of demand points in the region. Decision maker will need to choose an appropriate weighting for each demand point to ensure that the resulting solution reflects their overarching goals.

Targets are used to ensure that prioritizations sufficiently conserve each feature. Amount-based targets specify the minimum amount of habitat required for each feature to be adequately conserved (similar to those used in \texttt{Marxan}). Let $T_f$ denote the amount of area or habitat that needs to be preserved for each feature $f \in F$. Space-based targets specify the minimum proportion of variation in the demand points that needs to be secured for each feature to be sufficiently conserved. Space-based targets are expressed as proportions---instead of a sum of weighted distances---by scaling the sum of weighted distances between the demand points and the selected planning units in a solution relative to the distances between demand points and the demand points' centroid. This scaling is conceptually similar to that used in calculating the R$^2$ statistic for _K_-means cluster analyses from the within sums of squares and total sums of squares [page 106; @r491]. Let $\tau_{fs}$ denote the space-based targets for feature $f \in F$ and attribute space $s \in S$. The control variables for the unreliable formulation are the $B$, $T_{s}$, and $\tau_{fs}$ variables.

@@

Each demand point $i \in I$ for feature $f \in F$ and space $s \in S$ is represented by a single selected planning unit (ie. a $j \in J$ where $X_j = 1$). The degree to which a demand point $i$ is represented by a planning unit $j$ is determined by the distance between them ($d_{fsij}$). Generally---unless near zero space targets are used so that the problem is effectively unconstrained by the target---demand points are represented by their closest selected planning units. In poorer quality solutions, demand points will be represented by planning units that are further away from them. As a consequence, the sum of the weighted distances between all of the demand demand points and the planning units used to represent them will be larger, and so, the solution will capture less of the variation described by the demand points.

@@

The objective function (eqn 3a) determines the utility of a given prioritization: a combination of the total cost of a prioritization and how fragmented it is. Constraints (eqn 3b) ensure that all the amount-based targets are met. Constraints (eqn 3c) ensure that all the space-based targets are met for each feature and each attribute space. By treating targets as constraints, rather than including them in the objective function, all feasible solutions will fulfill the targets. For each feature and attribute space, the total weighted distance between the demand points and their closest selected planning units is calculated ($\sum_{i=0}^{I-1} \sum_{j=0}^{J-1} \lambda_{fsi} {d_{fsij}}^{2} Y_{fsij}$). This total weighted distance is then scaled by the inherent variation in the demand points ($\sum_{i=0}^{I-1} \lambda_{fsi} {\delta_{fsi}}^{2}$). The resulting fraction is used to calculate a proportion conceptually similar to the $R^2$ statistic used in $k$-means clustering analysis. The constraints (eqn 3c) ensure that the proportion of variation in the demand points secured in the solution must be equal to or greater than the space-based target. Constraints (eqn 3d) ensure that only one planning unit is assigned to each demand point. Constraints (eqn 3e) ensure that demand points are only assigned to selected planning units. Constraints (eqn 3f) ensure that the $X$ and $Y$ variables are binary.

## Optimization
Although the reserve selection problems presented here are non-linear (see Appendix S1 for the reliable formulation), they can be linearized using methods described by Beyer _et al._ [-@r426] and Cui _et al._ [@r16]. The \texttt{raptr} R package provides functions to express conservation planning data as linearized versions of the optimization problems and solve them using the commercial \texttt{Gurobi} software suite (\url{www.gurobi.com}). Presently, academics can obtain a special license for \texttt{Gurobi} at no cost.

@@

We generated four solutions for each species. First, to represent solutions generated using conventional methods (eg. \texttt{Marxan}), we generated solutions using `r sim_parameters$amount_target*100` % amount targets. Second, to show how the addition of space targets can affect solutions, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and `r (sim_parameters$space_target*100)` % space targets to capture the geographic spread of each species. Third, to represent solutions generated using conventional planning methods that penalize for fragmentation, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets and a boundary length modifier of `r sim_parameters$blm`. Finally, to illustrate the combined effects of using amount and space targets as well as fragmentation penalties, we generated solutions using `r (sim_parameters$amount_target*100)` % amount targets, `r (sim_parameters$space_target*100)` % space targets, and boundary length modifiers of `r sim_parameters$blm`.

@@

The solution generated for the uniformly distributed species using an amount target prioritized `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` planning units (Fig. 3a). For this scenario, all solutions containing `r dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$n` units are optimal because this species has an equal chance of occurring in any given planning unit. Although it may seem odd that the solution prioritized planning units along the southern end of the study area, this behavior occurred because there were many optimal solutions and the solver---without any other criteria to compare solutions---returned a solution that reflected the order that the data was encoded in the optimization problem. As we can see in later examples, this behavior only manifests when there are many optimal solutions. It is unlikely that conservation planners would encounter this behavior when using real-world data, and if they did, it would suggest that they need to obtain more data to inform the reserve selection process.

The solution generated using just amount targets only captured a small proportion of the variation in the geographic attribute space (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled). The coverage was so poor that this "proportion" was negative because the solution captured a less representative sample than a solution containing one planning unit in the center of the species' distribution. In other words, the total distance between the demand points and the 20 selected planning units was larger than the total distance between the demand points and a single planning unit in the center of the study area. Similarly, $R^2$ statistics can be have negative values when they are describing models that perform worse than a null model.

By explicitly targeting a representative sample of the species' geographic spread, we obtained a solution that captured the geographic spread of the uniform species (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3b). Although this solution secured an adequate amount of habitat and a representative sample of the species' geographic spread, this solution was highly fragmented. However, by penalizing fragmentation using a boundary length modifier, we were able to obtain a well connected solution that met all of the objectives (`r round(dplyr::filter(sim_spp_results, species == "Uniform\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig. 3d). This solution prioritized the same number of planning units as the previous solutions even though it is far superior. As we can see, under the simplest of circumstances, reserve selection methods may not yield solutions that secure a representative sample of features unless constraints are used to guarantee this property.

@@

The solution generated for the normally distributed species using just an amount target prioritized planning units in the species' core distribution (Fig. 3e). Although this strategy may seem cost-effective, simply conserving the places where this species is most likely to be found is a poor strategy for securing a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled) because it did not protect any peripheral habitat. By using amount and space targets, we obtained a solution that conserved an adequate amount of habitat and also secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount &\nspace targets")$space_held, 2)` % sampled; Fig. 3f). Similar to the solutions for the uniformly distributed species, this solution was highly fragmented and we were able to obtain a better connected solution by specifying fragmentation penalties (`r round(dplyr::filter(sim_spp_results, species == "Normal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig. 3h). However, unlike the solutions for the uniformly distributed species, the solution generated using an amount target, a space target, and fragmentation penalties required more planning units than the other solutions. These results suggest that solutions may need to prioritize more planning units to meet additional conservation objectives.

@@

The solution generated for the bimodally distributed species using just an amount target conserved individuals belonging to one of the two populations (Fig. 3i). As a consequence, this solution did not secure a representative sample of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount\ntargets")$space_held, 2)` % sampled). The addition of fragmentation penalties exacerbated this issue, and resulted in a solution that sampled even less of the species' geographic spread (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount target\nand BLM")$space_held, 2)` % sampled; Fig. 3k). However, this issue was resolved by using a space target  to obtain a solution that secured both populations (`r round(dplyr::filter(sim_spp_results, species == "Bimodal\nspecies", prioritisation == "Amount & space\ntargets and BLM")$space_held, 2)` % sampled; Fig. 3l). This finding suggests that species with large intra-specific variation could benefit the most from prioritizations generated using space-based targets.

## CASE STUDY 1
### Methods
We investigated how space-based targets can be used in a multi-species planning context to generate a prioritization that sufficiently preserves the species' realized niches. By preserving the populations in suitable habitats with different environmental conditions, conservation planners can preserve the species' adaptive landscape and foster resilience to environmental change [@r63]. We selected Queensland, Australia as the study area, and used an equal area coordinate system for geospatial analyses (Australian Albers GDA94; \texttt{EPSG:3577}). To represent planning units for this case-study, we generated a `r cs1_parameters$pu_size/1000` $\times$ `r cs1_parameters$pu_size/1000` km$^2$ grid and clipped the square polygons to the state boundary. We obtained data for 19 bioclimatic variables across the region [at $30^{\prime \prime}$ resolution from \url{www.worldclim.org}; @r33] and subjected them to a principal components analysis (using ArcMap 10.3.1). We used scores from the first `r as.character(english::as.english(cs1_parameters$number_pc_layers))` principal components to characterize the environmental variation across the study area (explaining `r round(cs1_pca_data[[4]][cs1_parameters$number_pc_layers], 1)` % of the total variation; Fig. 4).
